Journey: A Study of Scalabilty of Memory Management in Linux

### 2017-05-01 paper.tex
Working on kswapd section

TODO: in progress... should be ready for final product...

@@

TODO

>>>>

Figure (TODO) shows the total number of jiffies spent running kswapd as the
number of operations performed increases. Figure (TODO) shows the same for
khugepaged. Notice that one jiffy corresponds to 10ms of runtime.

We find that khugepaged does not take significant computational power. (TODO:
more) However, our workloads also are not conducive to allowing khugepaged to
merge pages, and it apparently does not choose to merge 2MB pages into 1GB
pages.

Likewise, we find that kswapd does not take up a significant percentage of
runtime compared with the applcation. The behavior of kswapd depends heavily on
the workload and on whether transparent hugepages are enabled. Workloads that do
not exhibit memory pressure do not incur kswapd overhead, as shown by the
\texttt{rand} benchmark with hugepages off.

As expected, the \texttt{rand} benchmark with hugepages on incurs the highest
total overhead since it begins swapping very quickly. For this benchmark we
observe an overhead of roughly 0.0002 jiffies per page memory touched. Figure
also shows that when huge pages are enabled, kswapd runs for longer periods of
time with longer sleeps in between. We suspect that kswapd swaps out several
large pages at a time, which frees more memory at once than free normal 4kB
pages, but due to time constraints, we have not verified this.

In contrast, the \texttt{cont} benchmark with hugepages disabled uses
computational time at roughly 0.0572 jiffies per page of memory touched.
Moreover, kswapd runs more frequently and for shorter periods of time. We
suspect that because pages are smaller, kswapd must run more often to reclaim
the same amount of memory as with large pages.

Notice also that the \texttt{cont} benchmark with hugepages enabled spends very
little time in kswapd. This is somewhat surprising given the amount of time
spent for the \texttt{rand} benchmark. We suspect that kswapd runs once and
choose a large number of pages to swap out. Subsequently, both the workload and
kswapd wait for I/O. This seems consistent with our previous results, in which
some operations take on the order of seconds to complete.

@@

kswapd and the swapping system seem to have some interesting pathological cases
that are worth exploring further.

We expect that this number increases with the size of physical memory
because kswapd will have more pages to scan, but due to time constraints, we
have not verified this.

TODO: anything else?
